Group-level homophily in metapopulation social networks can optimize the diffusion of innovations

### 2024-09-12 paper/main.tex
trophic dynamics working, figures made for initial writeup

\begin{abstract}
  \noindent
  In the Greater Mara, Kenya, across the past few decades, conservationists and
  policymakers have struggled to develop policies and practices that are robust to
  predictable and unpredictable impediments, and effectively optimize precious
  political, legal, and enforcement resources for biodiversity conservation. Indeed,
  biodiversity conservation worldwide struggles to meet critical targets, including
  the United Nations' Sustainable Development Goals 14 and 15, to protect, restore,
  and promote sustainable use of marine (Goal 14) and terrestrial ecosystems (Goal
  15).  We suggest that this is because any decades-long policy-based course of
  action in socio-ecological systems for is subject to uncertainty, stochasticity,
  and path-dependence of outcomes characteristic of the socio-ecological systems in
  which the policies. There are predictable impediments to biodiversity: these
  include personal-level cognitive biases or incentives of voters and policymakers
  such as the temptation of lucrative poaching or illegal grazing.  Another
  predictable impediment is if social networks are not sufficiently structured,
  biodiversity-protecting behaviors may fail to percolate through relevant
  populations via interpersonal observation and social. These impediments are indeed
  \emph{predictable} in some sense, but only in contexts that meet some set of
  theoretical assumptions and conditions, with observations necessarily limited to a
  finite number of experimental laboratory or real-world settings. We clearly cannot
  perform real-world experiments on the Mara, or any other biodiversity preserve,
  because we do not get more than one, and it must be saved now. Here we develop a
  strategy for anticipating a range of possible outcomes for biodiversity
  interventions by simulating the complex local-, individual-, and relatively
  shorter time-scale dynamics of stakeholder decision-making and trophic networks
  that are affected by regional, state, or national policy. For instance, we
  demonstrate that a push to incorporate more rangeland into conservancies can tip
  unincorporated pastoralists to build fences around their landholdings, which may
  eventually limit wild herbivore mobility by encircling herbivores in the reserve,
  which can become deadly. In one tragic example, in the case of a sudden severe
  drought, many landholders had been fencing their land over the course of many
  years, which had not been a problem since water had been plentiful. When the
  drought hit, wildebeest needed to migrate for water, but were unable to due to
  fence construction that only became an obvious problem after that tragic event.
  We explain how agent-based modeling can represent local dynamics, while these
  dynamics are constrained by national- and regional-level policy. These policies 
  are aimed at optimizing biodiversity outcomes over longer time scales, 
  which could be calculated, for example obtained through dynamic
  optimization that accounts for actor and stakeholder incentives. More local-,
  shorter time-scale predictions can be used to update optimal policies, which
  in turn generate new smaller-scale simulation outcomes, and so on so that
  the models at each scale contribute to a strategy for implementing long-term
  \emph{and} adaptive policy to protect biodiversity.
  \mt{Just an idea, haven't done this; I would very much welcome
  ideas on a Kenyan-focused or -inspired idea for a proof of concept like this.}
\end{abstract}

@@

@@

@@

\begin{itemize}
  \item
    Landholders (Table~\ref{tab:landholders})
  \item 
    Trophic network composed of five ``species'': insect, bird, herbivore,
    carnivore, and livestock. Of course livestock are a type of herbivore, but
    we need to separate them so that if they are being hunted by carnivores,
    for example, this can factor in to landowner decision making. Non-humans
    have a location, ``energy'' they have to move or give to those who consume
    them. Livestock are furthermore connected to a landholder, who own them and
    the land on which they roam.
\end{itemize}

@@

\begin{table}[h]
  \caption{\textbf{Non-human animal variables.} Landholders are allotted a plot of
  land with livestock and can choose to fence or not.}
  \label{tab:landholders}
  \begin{tabular}{cl} \toprule
    variable & description \\ \midrule  
    \textsc{position} & Spatial location of animal \\
    $\Delta$\textsc{energy} & Energy used to live for one time step \\
    \textsc{energy} & Amount of energy animal has left; animal dies when $<0$ \\
    \textsc{species} & Animal species: insect, bird, herbivore, carnivore, or
    livestock \\
    \textsc{landholder} & Pointer to landholder if animal is livestock; \texttt{NULL}
    otherwise \\
    \bottomrule
  \end{tabular} 
\end{table}

@@

\begin{enumerate}
  \item 
    Initialize 
\end{enumerate}

@@

>>>>

\begin{abstract}
  \noindent
  In the Greater Mara, Kenya, across the past few decades, conservationists and
  policymakers have struggled to develop policies and practices that are robust to
  predictable and unpredictable impediments, and effectively optimize precious
  political, legal, and enforcement resources for biodiversity conservation.
  This is one of many situations worldwide where 
  biodiversity conservation struggles to meet critical targets, including
  the United Nations' Sustainable Development Goal 15, to protect, restore,
  and promote sustainable use of terrestrial ecosystems (Goal
  15).  We suggest that this is because any decades-long policy-based course of
  action is subject to uncertainty, stochasticity,
  and path-dependence of outcomes characteristic of the socio-ecological systems in
  which the policies. We propose a multi-level modeling approach that optimizes policies 
  on the time scale of 
  about one century and spatial scale of the Greater Mara Ecosystem at a higher level,
  and at a lower level models landholder social learning and decision making and
  local trophic networks events over time scales of about ten years and
  spatial scales of ~100 square kilometers. 
  We believe these judicious simplifications of the socio-ecological system and incorporation
  in a mechanistic model may be used to design interventions to reduce fencing, 
  which is a known threat to biodiversity, especially in the Greater Mara.
\end{abstract}

@@

There are predictable impediments to biodiversity: these
include personal-level cognitive biases or incentives of voters and policymakers
such as the temptation of lucrative poaching or illegal grazing.  Another
predictable impediment is if social networks are not sufficiently structured,
biodiversity-protecting behaviors may fail to percolate through relevant
populations via interpersonal observation and social. These impediments are indeed
\emph{predictable} in some sense, but only in contexts that meet some set of
theoretical assumptions and conditions, with observations necessarily limited to a
finite number of experimental laboratory or real-world settings.

We clearly cannot
perform real-world experiments on the Mara, or any other biodiversity preserve,
because we do not get more than one, and it must be saved now. Therefore we 
develop a dual-pronged mechanistic modeling approach, where one prong attempts
to optimize long-term outcomes on the scale of one century, 
using optimal planning theory. The other approach
is more local in time and space, using agent-based models to represent 
roughly one decade of decision-making, climate variance, trophic networks, and
fencing decisions. Such mechanistic modeling enables us to predict how the past or future may have been or
could be different depending on contextual legal, environmental, and behavioral
factors~\cite{Turner2022}. Here we develop a
strategy for anticipating a range of possible outcomes for biodiversity
interventions by simulating the complex local-, individual-, and relatively
shorter time-scale dynamics of stakeholder decision-making and trophic networks
that are affected by regional, state, or national policy.

For instance, we may use this modeling approach to
demonstrate that a push to incorporate more rangeland into conservancies can tip
unincorporated pastoralists to build fences around their landholdings, which may
eventually limit wild herbivore mobility by encircling herbivores in the reserve,
which can become deadly. In one tragic example, in the case of a sudden severe
drought, many landholders had been fencing their land over the course of many
years, which had not been a problem since water had been plentiful. When the
drought hit, wildebeest needed to migrate for water, but were unable to due to
fence construction that only became an obvious problem after that tragic event.
We explain how agent-based modeling can represent local dynamics, while these
dynamics are constrained by national- and regional-level policy. These policies 
are aimed at optimizing biodiversity outcomes over longer time scales, 
which could be calculated, for example obtained through dynamic
optimization that accounts for actor and stakeholder incentives. More local-,
shorter time-scale predictions can be used to update optimal policies, which
in turn generate new smaller-scale simulation outcomes, and so on so that
the models at each scale contribute to a strategy for implementing long-term
\emph{and} adaptive policy to protect biodiversity.

In the rest of this modeling sketch I explain an agent-based model for more
local, shorter time-scale modeling of climate variance, fencing decisions and
trophic networks.

@@

To understand how fencing decisions may increase the risk of biodiversity loss,
we developed an agent-based model representing vegetation, a simple trophic network
(Figure~\ref{fig:TrophicNetwork}),
landholders (Table~\ref{tab:landholders}), their livestock, and, most importantly, 
their fences.

@@

\subsubsection{Landholders}

\begin{itemize}
  \item
    Landholders (Table~\ref{tab:landholders})
  \item 
\end{itemize}

@@

\subsubsection{Vegetation and trophic network}

Trophic network composed of five ``species'': insect, bird, herbivore,
carnivore, and livestock. Of course livestock are a type of herbivore, but
we need to separate them so that if they are being hunted by carnivores,
for example, this can factor in to landowner decision making.

\begin{figure}
  \caption{\textbf{Trophic network illustration.}}
  \centering
    \includegraphics[width=0.375\textwidth]{Figures/TrophicNetwork.pdf}
  \label{fig:TrophicNetwork}
\end{figure}

Non-humans
have a location, ``energy'' they have to move or give to those who consume
them. Livestock are furthermore connected to a landholder, who own them and
the land on which they roam.

\begin{table}[h]
  \caption{\textbf{Non-human animal variables.} All non-human species have
  the share a set of attributes. Predators have an additional attribute, called
\textsc{catch\_radius}, specifying how close prey must be in order to catch it.}

  \label{tab:nonhuman}
  \begin{tabular}{cl} \toprule
    variable & description \\ \midrule  
    \textsc{position} & Spatial location of animal \\
    $\Delta$\textsc{energy} & Energy used to live for one time step \\
    \textsc{energy} & Amount of energy animal has left; animal dies when $<0$ \\
    \textsc{species} & Animal species: insect, bird, herbivore, carnivore, or
    livestock \\
    \textsc{landholder} & Pointer to landholder if animal is livestock; \texttt{NULL}
    otherwise \\
    \textsc{catch\_radius} & Distance away for which a prey animal can be caught; \texttt{NULL} for non-predators  \\
    \bottomrule
  \end{tabular} 
\end{table}

@@

@@

\begin{figure}
  \caption{\textbf{Spatial time series of fence construction.} Fences (pink
  polygon overlays) are constructed over time in response to carnivore
  predation of livestock and encroachment by other herbivores. Vegetation is
  shown in green, with darker cells indicating less vegetation due to consumption
  by livestock and other herbivores. Time $t$ is in months.}
  \centering
  \includegraphics[width=0.9\textwidth]{Figures/fencing_series_10y.pdf}
  \label{fig:fence_series}
\end{figure}

\begin{figure}
  \caption{\textbf{Biodiversity prevalence dynamics.}}
  \centering
    \includegraphics[width=0.7\textwidth]{Figures/prevalence_dynamics.pdf}
  \label{fig:prevalence_dynamics}
\end{figure}
